Historic collections as a tool for assessing the global pollination crisis

### 2018-06-14 revision/ms.Rmd
have added a new reference list (references2) and made some minor edits on the manuscript (in revision/ms)

---
title: Historic collections as a tool for assessing the global pollination crisis
author: "Bartomeus, I.^1^, Stavert, J.R.^2^, Ward, D.^2,3^, and Aguado, O.^4^"
csl: data/philb_style.xml
output:
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
bibliography: data/references.xml
---

@@

One of the main barriers to identifying long-term pollinator population trends is that pollinators are incredibly taxonomically diverse, including bees, flies, butterflies, beetles, birds, bats and lizards [@Ollerton2017]. Additionally, many pollinators are highly mobile, short-lived and small, which makes monitoring their populations difficult. Bees are generally regarded as the most important pollinator group due to their abundance, pollination efficiency and widespread distribution [@Winfree2011]. However, bees are diverse, with more than 20,000 species currently described worldwide, and often require expert taxonomists for identification. Furthermore, the worldwide uneven distribution of research studies has resulted in geographical biases in bee decline research [@Martin2012], as well as taxonomic biases toward species that are easier to identify, such as bumblebees [@Archer2014; @DePalma2016].

@@

At a global scale, current evidence of pollinator declines is highly limited with most data restricted to the US and Europe. It is unsurprising that studies on pollinator declines are biased towards developed western countries, which have also been subject to extensive anthropogenic disturbance. For example, in the UK and the Netherlands, a citizen science based study using both observations and museum collection data detected strong richness declines for bees, hoverflies and flowering plants [@Biesmeijer2006]. In the Netherlands, museum data have also revealed coupled plant and pollinator declines [@Scheper2014]. Specifically, bee species with the strongest host plant preferences (i.e., specialists) displayed the strongest declines and thus, were most threatened with extinction. However, it is important to note that even for these two countries, local estimates of pollinator richness are biased toward large cities and regions dominated by agriculture, and thus lack data for less human-developed areas [jamie check wording of the "human-developed"](). Further exploration of this dataset revealed that for declining pollinator taxa, the species richness declining trend has attenuated in recent decades [@Carvalheiro2013].

Although studies of local native pollinator communities often detect richness declines, regional native richness may remain relatively stable. For example, regional estimates for bee species richness changes in the eastern US show moderate declines [@Bartomeus2013a] and very few regional extinctions [@Colla2012]. This is a pattern also detected in the UK, where relatively few regional bee extinctions has been reported [@Ollerton2014]. These regional findings are in stark contrast with the widespread local extinctions reported in local studies. For example, Burkle et al. [@Burkle2013] compared historical observations of bee species’ occurrences in a large forested ecosystem with remaining forest remnants and reports several local extinctions. However, it is important to note that there is strong concordance between local extinctions and regional declines [@Bartomeus2013c], suggesting that local extinctions are indicators of regional population declines.

Reported declines for bumblebees are the most severe of all pollinator taxa. For example, declines of up to 18% in local bumblebee richness have been reported for Belgium and the Netherlands [@Carvalheiro2013]. In other parts of Europe, local richness declines range from 5% in Great Britain [@Carvalheiro2013] to 42% in Denmark [@Dupont2011] and 70 % of bumblebee species are clasified as threatened or with declining population trends [@Nieto2014]. In the USA, reported bumble decline are also severe with estimates ranging between 25% [@Grixti2009] and 30% [@Bartomeus2013a]. However, studies on species richness changes for other pollinator taxa are both scarce and geographically restricted. For butterflies, the only evidence of richness declines comes from Europe. Butterfly species richness has declined substantially in the Netherlands and Belgium since the 1950’s, although declines in Great Britain have been less severe [@Carvalheiro2013]. In Belgium, another study [@Maes2001] found that richness declines have been severe (approximately 30%), although this study assessed richness changes over a longer time period (early 1900’s to 2000) compared to [@Carvalheiro2013] (1950-69 vs. 1970-80 and 1970-89 vs. 1990-2009). In parts of Germany, up to 70% declines in local butterfly richness have been reported [@Wenzel2006]. Compared with other insect pollinator taxa, there are very few studies on hoverfly species richness changes, which are all restricted to Europe. In Belgium, Great Britain and the Netherlands, hoverfly richness changes have been modest [@Carvalheiro2013]. In the Netherlands, moderate increases in hoverfly species richness have been shown, whereas in Great Britain no significant directional changes were detected [@Biesmeijer2006]. Furthermore, directionality (richness increase or decrease) varies depending on the time period assessed. For example, hoverfly richness decreased in Belgium by approximately 6% from 1950-69 to 1970-80, but increased by approximately 10% between 1970-89 and 1990-2009 [@Carvalheiro2013].

For ilustrative purposes, we mapped the findings of these studies in Figure 1 to show the strong contrast between bee species richness worldwide, whith bee diversity hotspots in mediterranean countries, against the paucity of countries for which we have any local or regional data on bee, syrphid or butterfly declines (raw data: Sup Mat 1). Except for outside of Europe and the US and for non-insect taxa, there are very few or no studies on pollinator declines using historical records. However, there are species-specific examples of historical losses from different parts of the world (e.g., *Bombus dalbhomi*; @Morales2013).

@@

Where digitization has been completed, the data provide a rich source of information, allowing assessment of the current status and long-term trends of pollinator populations [e.g., @Bartomeus2013a; @Scheper2014; @Young2017]. This is despite the fact that museum collections often have a number of biases, including unknown sampling effort, personal interests of collectors and the curatorial techniques used. For example, collectors tend to target rare or unusual over common taxa, discard damaged individuals or only accession a certain number of individuals. In addition, collections are often made opportunistically, which may lead to spatial bias where difficult to access areas are under-sampled or conversely, where samples are biased towards easily accessed locations (e.g., towns/cities and/or roadsides). Further, museum collection data can only be used to determine where species are present and not where they are absent. However, given adequate sample sizes and appropriate statistical techniques, most biases can be accounted for [e.g., @Bartomeus2013a; @Keil2011; @Pearce2006].

@@

GBIF (https://www.gbif.org/) is a central repository for global species occurrence data. Much of these data come from museums, private collections and government research institutes, but several other sources are also integrated. In combination with the popular statistical language R [@RCoreTeam2017], GBIF can be directly queried into your computer [@Chamberlain2014] and data availability can be checked for the region of interest. Focusing on bee taxa, we show here the number of modern and historic bee records currently available for different countries (Figure 2a, Sup mat 2). We selected thirty-seven countries that count with more than 1800 records in each time period, making these data potentially analyzable without further data collection effort (see Figure 2b and c for an initial exploration). We selected this arbitrary threshold as a compromise between sample size and number of countries that can be included in a general analysis. However, a proper analysis of this dataset and the minimum sample size needed per country would need a careful look at the data characteristics as we detail later on with two examples (Spain and New Zealand). Most countries fall short in one or both axes. For example, a variety of countries located in different continents such as Switzerland, Sri Lanka, Nicaragua or Zimbawe have a decent number of recent records, but lack historical collections and researchers should prioritize the digitalization of old material before embarking on data analyses. It is also important to note that historical records are not always vouchered in local museums (i.e., many European and US museums contain large collections of pollinators from other countries). In contrast, more than 192 countries have less than 1000 records for both time periods, making them poor candidates for analysing long-term pollinator population trends. We focus on bees, but similar exploratory analyses can easily be conducted for other taxa.

@@

As stated above, once historical datasets are made available, researchers must identify any potential biases. We explore this process with two contrasting dataset examples (Spain and New Zealand). In the Spanish dataset, most of the data comes from a few specific locations and was collected by a few specific teams. Hence the geographical coverage is not representative. Even worse, historical and modern collections do not overlap spatially, making any inference imposible to interpret. In this case, contacted the original collectors of the historical data to define their sampling protocols. We then resurveyed the same sites (30 years after the original surveys) using the same sampling protocols. In contrast, the New Zealand dataset includes a wide suite of collectors and collection locations but shows no obvious biases in geographical and taxonomic coverage through time. We complemented gbif data with further museum collections for bees and flies and analyze the regional richness changes through time. For these two case studies, we provide annotated R scripts as examples of analysis for different dataset types (Sup Mat 3). These different analytical approaches allow us to reveal long-term trends in pollinator populations for regions with contrasting sampling histories. We hope this resource will encourage researchers to analyse data for regions where current information on pollinator declines is lacking.

@@

In contrast to Spain, New Zealand is an isolated oceanic archipelago, with a distinctive pollinator biota and a unique history of human occupation. Much of New Zealand’s pollinator fauna is also relatively depauperate. For example, New Zealand has only 27 native bee species [@Donovan2007], which is a fraction of nearby Australia’s c. 1600 species [@Ascher2015]. However, New Zealand has a surprisingly high diversity of flies (Diptera), which are important pollinators in many ecosystems [@Newstrom2005]. Thus, New Zealand provides a unique system to study long-term changes in pollinator communities, and is unlike continental Europe and the US, which have been the focus of an overwhelming majority of pollinator decline studies.

In global terms, human colonisation of New Zealand was relatively recent (c. 740 y) [@Wilmshurst2008]. Before human arrival, New Zealand was predominately forested, but has since been dramatically altered by people. We can use museum records to identify trends in pollinator communities during New Zealand’s more recent history.

We used New Zealand bee collection records gathered from multiple sources, including university, research institute, museum and private collections. Collection records from the New Zealand Arthropod Collection (NZAC) are freely available online (https://scd.landcareresearch.co.nz/). Fly pollinator data was obtained from three participating New Zealand museums and covers two families (Calliphoridae and Syrphidae) that contain important fly pollinators. Collections for the bee and fly datasets span over 100 years (early 1900s to late 2000s).

@@

Second, to determine if the probability of finding a species in the collection changed over time, we used a general linear model with a binomial distribution and a logit link. For species that showed overdispersion, we used a quasi-binomial distribution. Further, we only included species in this analysis for which we had 30 or more records. Preliminary analysis using random subsampling of specimens from species with large sample size indicated that below 30 records the trends become too variable. To account for differences in sampling effort between years, we weighted each year by the total number of samples collected that year.

@@

p <- ggplot(est.all, aes(Var1, estimate_frq, group=taxa))
p <- p + xlab(NULL) + ylab("Change in relative occurance")
p <- p + facet_wrap(~taxa,nrow=2,ncol=1,scales = "free_x")
p <- p + geom_point(aes(colour=change_in_occurance), size = 2)
p <- p + geom_errorbar(aes(ymin=estimate_frq-SE_frq, ymax=estimate_frq+SE_frq, colour=change_in_occurance), width = 0)
p <- p + geom_hline(aes(yintercept=0), linetype="dashed")
p <- p + geom_vline(data=filter(est.all, taxa=="Bee"), aes(xintercept=5.5), colour="grey")
p <- p + geom_vline(data=filter(est.all, taxa=="Fly"), aes(xintercept=7.5), colour="grey")
p <- p + geom_text(data=data.frame(Var1=3,estimate_frq=0.145,taxa="Bee",lab='Exotic'), aes(label=lab), parse=TRUE,colour="grey70", size=3)
p <- p + geom_text(data=data.frame(Var1=4,estimate_frq=0.145,taxa="Fly",lab='Exotic'), aes(label=lab), parse=TRUE,colour="grey70", size=3)
p <- p + geom_text(data=data.frame(Var1=16.5,estimate_frq=0.145,taxa="Bee",lab='Native'), aes(label=lab), parse=TRUE,colour="grey70", size=3)
p <- p + geom_text(data=data.frame(Var1=11,estimate_frq=0.145,taxa="Fly",lab='Native'), aes(label=lab), parse=TRUE,colour="grey70", size=3)
p <- p + theme(panel.grid.minor = element_blank(),
               panel.background = element_blank(),
               axis.line = element_line(colour = "black")) +
    theme(panel.border=element_rect(colour = "black", fill = "NA", size = 1)) +
    theme(axis.text.y=element_text(angle= 360, hjust = 0.5, vjust = 0.5, size =12),
          axis.title.y=element_text(size=18, vjust = 1),
          axis.text.x=element_text(face = "italic", angle= 45, hjust = 1, vjust = 1, size =7),
          axis.title.x=element_text(size=22, vjust = 1),
          axis.text=element_text(colour = "black"))+
    theme(strip.background = element_rect(colour="NA", fill=NA),
          strip.text = element_text(size=14))
p <- p + theme(axis.title.y=element_text(margin=margin(0,20,0,0)))
p <- p + scale_colour_brewer(palette = "Dark2")
p$labels$colour <- "Change in occurance"
p

@@

A recent study found that more than 90% of the papers investigating pollinator responses to land-use change focused solely on richness and abundance descriptors [@Winfree2011]. But in addition to local (alpha) diversity and regional (gamma) diversity, researchers need to assess changes in turnover between sites (beta diversity). Environmental changes often result in a few “winner” species and many “loser” species [@Bartomeus2013a]. Identifying winners and losers is critical as the few winners are often exotic and represent a subset of traits that facilitate survival in highly modified environments [@Stavert2017]. These changes can have important effects for pollination of native plant species and crops [@Stavert2018].

In addition, digitalized museum specimen collections can provide much more information besides species occurrence records, given that such information is recorded when digitizing collections. This is particularly important for identifying mechanisms of decline and adaptation. For example, recording the date of collection is particularly important for tracking of phenological advances congruent with contemporary climate change [@Bartomeus2011]. In addition, pollinator specimen labels often include information about the host plant on which the specimen was collected. This information critical for understanding past and present species interactions [@Tylianakis2008]. Aside from this information, bee specimens often contain pollen loads trapped on hairs, from which past visitation events can be identified [@Kleijn2008]. Finally, museum specimens can be measured to track evolutionary changes by measuring the traits of specimen. This approach has been already used to investigate tongue length [@Miller-Struttmann2015] and body size [@Oliveira2016, @Renauld2016] [jamie, ref to add]() in response to climate and land-use change. Finally, plant specimens stored in plant herbaria can also contain indirect evidence of pollinator and pollination declines [60] which can link pollinator declines with its consequences for ecosystem functioning.

@@

Unleashing the power of museum collection data to answer pressing ecological and evolutionary questions is at our hands, but requires the coordinated effort of many actors. Using two case studies, we show that strong collaboration between museum curators and ecologists is key to understanding data and treating it appropriately. To advance our understanding of the global pollination crisis, researchers and curators must aim to digitize museum collection data and make it readily available in a format that is widely accessible. Centralization of regional and national museum collection data in existing global platforms, such as GBIF, would facilitate free and widespread access. However, datasets could also be stored in alternative webpages or database repositories (e.g., university and museum webpages or Dryad) providing they are thoroughly documented and easily retrieved and combined with other datasets using open science tools [e.g., @Dallas2016].

We must revolutionize the way that researchers collaborate with museums, in order to foster healthy bidirectional relationships. For example, ecological researchers collect massive amounts of specimens, but these are often inappropriately vouchered [@Packer2018; @Wheeler2003], rendering them less useful for future research. To improve this process, strong communication between museums and researchers is required. However, this can only be achieved with adequate funding and recognition that accurate data recording and long-term preservation are critical for research [@Ward2015].

To identify global trends in pollinator declines we require robust data, collected from diverse geographic regions. It is also crucial that these data are analysed appropriately. This requires researches to identify biases and to fill any taxonomic and geographic gaps where possible. We need to place increased emphasis on quantifying pollinator declines in regions outside of the US and Europe, and for pollinator groups other than bees. For the US and Europe, there have been few regional bee extinctions [@Ollerton2014; @Bartomeus2013a] but in disturbed ecosystems, declines are widespread [@DePalma2017; @Biesmeijer2006]. For most other pollinator taxa and regions throughout the world we know almost nothing. Moving forward, the first step for many taxa will be to identify and describe species [REF]. Only then can we begin to document pollinator declines.

>>>>

---
title: Historic collections as a tool for assessing the global pollination crisis
author: "Bartomeus, I.^1^, Stavert, J.R.^2^, Ward, D.^2,3^, and Aguado, O.^4^"
csl: data/philb_style.xml
output:
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
bibliography: data/references2.xml
---

@@

One of the main barriers to identifying long-term pollinator population trends is that pollinators are incredibly taxonomically diverse and include bees, flies, butterflies, beetles, birds, bats and lizards [@Ollerton2017]. Additionally, many pollinators are highly mobile, short-lived and small in size, which makes monitoring their populations difficult. Bees are generally regarded as the most important pollinator group due to their abundance, pollination efficiency and widespread distribution [@Winfree2011]. However, bees are diverse, with more than 20,000 species currently described worldwide, and often require expert taxonomists for identification. Furthermore, the uneven global distribution of research studies has resulted in geographical biases in bee decline research [@Martin2012], as well as taxonomic biases toward species that are easier to identify, such as bumblebees [@Archer2014; @DePalma2016].

@@

At a global scale, current evidence of pollinator declines is highly limited with most data restricted to the US and Europe. It is unsurprising that studies on pollinator declines are biased towards developed western countries, which have also been subject to extensive anthropogenic disturbance. For example, in the UK and the Netherlands, a citizen science based study using both observations and museum collection data detected strong richness declines for bees, hoverflies and flowering plants [@Biesmeijer2006]. In the Netherlands, museum data have also revealed coupled plant and pollinator declines [@Scheper2014]. Specifically, bee species with the strongest host plant preferences (i.e., specialists) displayed the strongest declines and thus, were most threatened with extinction. However, it is important to note that even for these two countries, local estimates of pollinator richness are biased toward large cities and regions dominated by agriculture, and thus lack data for areas subject to less human activity. Further exploration of this dataset revealed that for declining pollinator taxa, the declining species richness trend has attenuated in recent decades [@Carvalheiro2013].

Although studies of local native pollinator communities often detect richness declines, regional native richness may remain relatively stable. For example, regional estimates for bee species richness changes in the eastern US show moderate declines [@Bartomeus2013a] and very few regional extinctions [@Colla2012]. This is a pattern also detected in the UK, where relatively few regional bee extinctions has been reported [@Ollerton2014]. These regional findings contrast with the local extinctions reported in other studies. For example, Burkle et al. [@Burkle2013] compared historical observations of bee species’ occurrences in a large forested ecosystem with remaining forest remnants and reported several local extinctions. However, it is important to note that there is strong concordance between local extinctions and regional declines [@Bartomeus2013c], suggesting that local extinctions are indicators of regional population declines.

Reported declines for bumblebees are the most severe of all pollinator taxa. For example, declines of up to 18% in local bumblebee richness have been reported for Belgium and the Netherlands [@Carvalheiro2013]. In other parts of Europe, local richness declines range from 5% in Great Britain [@Carvalheiro2013] to 42% in Denmark [@Dupont2011] and 70 % of bumblebee species are clasified as threatened or with declining population trends [@Nieto2014]. In the USA, reported bumble declines are also severe with estimates ranging between 25% [@Grixti2009] and 30% [@Bartomeus2013a]. However, studies on species richness changes for other pollinator taxa are both scarce and geographically restricted. For butterflies, the only evidence of richness declines comes from Europe. Butterfly species richness has declined substantially in the Netherlands and Belgium since the 1950’s, although declines in Great Britain have been less severe [@Carvalheiro2013]. In Belgium, another study [@Maes2001] found that butterfly richness declines have been severe (approximately 30%), although this study assessed richness changes over a longer time period (early 1900’s to 2000) compared to [@Carvalheiro2013] (1950-69 vs. 1970-80 and 1970-89 vs. 1990-2009). In parts of Germany, up to 70% declines in local butterfly richness have been reported [@Wenzel2006]. Compared with other insect pollinator taxa, there are very few studies on hoverfly species richness changes, of which are all restricted to Europe. Specifically, in Belgium, Great Britain and the Netherlands, hoverfly richness changes have been modest [@Carvalheiro2013]. In the Netherlands, moderate increases in hoverfly species richness have been shown, whereas in Great Britain no significant directional changes were detected [@Biesmeijer2006]. Furthermore, directionality (richness increase or decrease) varies depending on the time period assessed. For example, hoverfly richness decreased in Belgium by approximately 6% from 1950-69 to 1970-80, but increased by approximately 10% between 1970-89 and 1990-2009 [@Carvalheiro2013].

For illustrative purposes, we mapped the findings of these studies (Figure 1) to show the large differences in bee species richness worldwide, with bee diversity hotspots in mediterranean countries, against the paucity of countries for which we have any local or regional data on bee, syrphid or butterfly declines (raw data: Sup Mat 1). Except for outside of Europe and the US and for non-insect taxa, there are very few or no studies on pollinator declines using historical records. However, there are species-specific examples of historical losses from different parts of the world (e.g., *Bombus dalbhomi*; @Morales2013).

@@

Where digitization has been completed, the data provide a rich source of information, allowing assessment of the current status and long-term trends of pollinator populations [e.g., @Bartomeus2013a; @Scheper2014; @Young2017]. This is despite the fact that museum collections often have a number of biases, including unknown sampling effort, personal interests of collectors and the curatorial techniques used. For example, collectors tend to target rare or unusual over common taxa, discard damaged individuals or only accession a certain number of individuals. In addition, collections are often made opportunistically, which may lead to spatial bias where difficult to access areas are under-sampled or conversely, where samples are biased towards easily accessed locations (e.g., towns, cities and roadsides). Further, museum collection data can only be used to determine where species are present, not where they are absent. However, given adequate sample sizes and appropriate statistical techniques, most biases can be accounted for [e.g., @Bartomeus2013a; @Keil2011; @Pearce2006].

@@

GBIF (https://www.gbif.org/) is a central repository for global species occurrence data. Much of these data come from museums, private collections and government research institutes, but several other sources are also integrated. In combination with the popular statistical language R [@RCoreTeam2017], GBIF can be directly queried into your computer [@Chamberlain2014] and data availability can be checked for the region of interest. Focusing on bee taxa, we show here the number of modern and historic bee records currently available for different countries (Figure 2a, Sup mat 2). We selected thirty-seven countries that have more than 1800 records in each time period, making these data potentially analyzable without further data collection effort (see Figure 2b and c for an initial exploration). We selected this arbitrary threshold as a compromise between sample size and the number of countries that can be included in a general analysis. However, proper analysis of this dataset and determining the minimum sample size needed per country would require careful inspection of the data, which we outline later with two examples (Spain and New Zealand). Most countries fall short in one or both axes. For example, a variety of countries located in different continents such as Switzerland, Sri Lanka, Nicaragua or Zimbawe have a decent number of recent records, but lack historical collections. Thus, researchers should prioritize the digitalization of old material before embarking on data analyses for these countries. It is also important to note that historical records are not always vouchered in local museums (i.e., many European and US museums contain large collections of pollinators from other countries). In contrast, more than 192 countries have less than 1000 records for both time periods, making them poor candidates for analysing long-term pollinator population trends. We focus on bees, but similar exploratory analyses can easily be conducted for other taxa.

@@

As stated above, once historical datasets are made available, researchers must identify any potential biases. We explore this process with two contrasting dataset examples (Spain and New Zealand). In the Spanish dataset, most of the data comes from a few specific locations and was collected by a few specific teams. Hence the geographical coverage is not representative. Even worse, historical and modern collections do not overlap spatially, making any inference impossible to interpret. In this case, we contacted the original collectors of the historical data to define their sampling protocols. We then resurveyed the same sites (30 years after the original surveys) using the same sampling protocols. In contrast, the New Zealand dataset includes a wide suite of collectors and collection locations but shows no obvious biases in geographical and taxonomic coverage through time. We complemented gbif data with further museum collections for bees and flies and analyze the regional richness changes through time. For these two case studies, we provide annotated R scripts as examples of analysis for different dataset types (Sup Mat 3). These different analytical approaches allow us to reveal long-term trends in pollinator populations for regions with contrasting sampling histories. We hope this resource will encourage researchers to analyse data for regions where current information on pollinator declines is lacking.

@@

In contrast to Spain, New Zealand is an isolated oceanic archipelago, with a distinctive pollinator biota and a unique history of human occupation. Much of New Zealand’s pollinator fauna is also relatively depauperate. For example, New Zealand has only 28 native bee species [@Donovan2007], which is a fraction of nearby Australia’s c. 1600 species [@Ascher2015]. However, New Zealand has a surprisingly high diversity of flies (Diptera), which are important pollinators in many ecosystems [@Newstrom2005]. Thus, New Zealand provides a unique system to study long-term changes in pollinator communities, and is unlike continental Europe and the US, which have been the focus of an overwhelming majority of pollinator decline studies.

In global terms, human colonisation of New Zealand was relatively recent (c. 740 y) [@Wilmshurst2008]. Before human arrival, New Zealand was predominately forested, but has since been dramatically altered by people. Therefore, we can use museum records to identify trends in pollinator communities during New Zealand’s more recent history.

We used New Zealand bee collection records gathered from multiple sources, including university, research institute, museum and private collections. Collection records from the New Zealand Arthropod Collection (NZAC) are freely available online (https://scd.landcareresearch.co.nz/). Fly pollinator data were obtained from three participating New Zealand museums and covers two families (Calliphoridae and Syrphidae) that contain important fly pollinators. Collections for the bee and fly datasets span over 100 years (early 1900s to late 2000s).

@@

Second, to determine if the probability of finding a species in the collection changed over time, we used a general linear model with a binomial distribution and a logit link. For species that showed overdispersion, we used a quasi-binomial distribution. Further, we only included species in this analysis for which we had 30 or more independent records [@Bartomeus2013a]. Preliminary analysis using random subsampling of specimens from species with large sample size indicated that below 30 records the trends become too variable. To account for differences in sampling effort between years, we weighted each year by the total number of samples collected that year.

@@

p <- ggplot(est.all, aes(Var1, estimate_frq, group=taxa))
p <- p + xlab(NULL) + ylab("Change in relative occurance")
p <- p + facet_wrap(~taxa,nrow=2,ncol=1,scales = "free_x")
p <- p + geom_point(aes(colour=change_in_occurance), size = 2)
p <- p + geom_errorbar(aes(ymin=estimate_frq-SE_frq, ymax=estimate_frq+SE_frq, colour=change_in_occurance), width = 0)
p <- p + geom_hline(aes(yintercept=0), linetype="dashed")
p <- p + geom_vline(data=filter(est.all, taxa=="Bee"), aes(xintercept=5.5), colour="grey")
p <- p + geom_vline(data=filter(est.all, taxa=="Fly"), aes(xintercept=7.5), colour="grey")
p <- p + geom_text(data=data.frame(Var1=3,estimate_frq=0.145,taxa="Bee",lab='Exotic'), aes(label=lab), parse=TRUE,colour="grey70", size=3)
p <- p + geom_text(data=data.frame(Var1=4,estimate_frq=0.145,taxa="Fly",lab='Exotic'), aes(label=lab), parse=TRUE,colour="grey70", size=3)
p <- p + geom_text(data=data.frame(Var1=16.5,estimate_frq=0.145,taxa="Bee",lab='Native'), aes(label=lab), parse=TRUE,colour="grey70", size=3)
p <- p + geom_text(data=data.frame(Var1=11,estimate_frq=0.145,taxa="Fly",lab='Native'), aes(label=lab), parse=TRUE,colour="grey70", size=3)
p <- p + theme(panel.grid.minor = element_blank(),
               panel.background = element_blank(),
               axis.line = element_line(colour = "black")) +
    theme(panel.border=element_rect(colour = "black", fill = "NA", size = 1)) +
    theme(axis.text.y=element_text(angle= 360, hjust = 0.5, vjust = 0.5, size =12),
          axis.title.y=element_text(size=18, vjust = 1),
          axis.text.x=element_text(face = "italic", angle= 45, hjust = 1, vjust = 1, size =8),
          axis.title.x=element_text(size=22, vjust = 1),
          axis.text=element_text(colour = "black"))+
    theme(strip.background = element_rect(colour="NA", fill=NA),
          strip.text = element_text(size=14))
p <- p + theme(axis.title.y=element_text(margin=margin(0,20,0,0)))
p <- p + theme(legend.text=element_text(size=12))
p <- p + theme(legend.title = element_blank())
p <- p + scale_colour_brewer(palette = "Dark2")
p$labels$colour <- "Change in occurance"
p

@@

A recent study found that more than 90% of the papers investigating pollinator responses to land-use change focused solely on richness and abundance descriptors [@Winfree2011]. However, in addition to local (alpha) diversity and regional (gamma) diversity, researchers need to assess changes in turnover between sites (beta diversity). Environmental changes often result in a few “winner” species and many “loser” species [@Bartomeus2013a]. Identifying winners and losers is critical as the few winners are often exotic and represent a subset of traits that facilitate survival in highly modified environments [@Stavert2017]. These changes can have important effects for pollination of native plant species and crops [@Stavert2018].

In addition, digitalized museum specimen collections can provide much more information besides species occurrence records, given that such information is recorded when digitizing collections. This is particularly important for identifying mechanisms of decline and adaptation. For example, recording the date of collection is particularly important for tracking of phenological advances congruent with contemporary climate change [@Bartomeus2011]. In addition, pollinator specimen labels often include information about the host plant on which the specimen was collected. This information critical for understanding past and present species interactions [@Tylianakis2008]. Aside from this information, bee specimens often contain pollen loads trapped on hairs, from which past visitation events can be identified [@Kleijn2008]. Finally, museum specimens can be measured to track evolutionary changes by measuring the traits of specimens. This approach has already been used to investigate tongue length [@Miller-Struttmann2015] and body size [@Oliveira2016, @Renauld2016] in response to climate and land-use change. Finally, plant specimens stored in plant herbaria may contain indirect evidence of pollination declines [60], thus linking pollinator declines with consequences for ecosystem functioning.

@@

Unleashing the power of museum collection data to answer pressing ecological and evolutionary questions is at our hands, but requires the coordinated effort of many actors. Using two case studies, we show that strong collaboration between museum curators and ecologists is key to understanding data and treating it appropriately. To advance our understanding of the global pollination crisis, researchers and curators must aim to digitize museum collection data and make it readily available in a format that is widely accessible. Centralization of regional and national museum collection data in existing global platforms, such as GBIF, would facilitate free and widespread access. However, datasets could also be stored in alternative webpages or database repositories (e.g., university and museum webpages or Dryad) providing they are thoroughly documented, and easy to retrieve and combined with other datasets using open science tools [e.g., @Dallas2016].

We must revolutionize the way that researchers collaborate with museums in order to foster healthy bidirectional relationships. For example, ecological researchers collect massive amounts of specimens, but these are often inappropriately vouchered [@Packer2018; @Wheeler2003], rendering them less useless for future research. To improve this process, strong communication between museums and researchers is required. However, this can only be achieved with adequate funding and recognition that accurate data recording and long-term preservation are critical for research [@Ward2015].

To identify global trends in pollinator declines we require robust data, collected from diverse geographic regions. It is also crucial that these data are analysed appropriately. This requires researches to identify biases and to fill any taxonomic and geographic gaps where possible. We need to place increased emphasis on quantifying pollinator declines in regions outside of the US and Europe, and for pollinator groups other than bees. For the US and Europe, there have been few regional bee extinctions [@Ollerton2014; @Bartomeus2013a] but in disturbed ecosystems, declines are widespread [@DePalma2017; @Biesmeijer2006]. For most other pollinator taxa and regions throughout the world, we know almost nothing. Moving forward, the first step for many taxa will be to identify and describe species [REF]. Only then can we begin to document pollinator declines.
